---
title: "Practical Machine Learning Course Project"
author: "E.T. Moseley"
date: "January 28, 2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
#Executive Summary

how you built your model
how you used cross validation
what you think the expected out of sample error is
why you made the choices you did

In the aforementioned study, six participants participated in a dumbell lifting exercise five different ways. 

The five ways, as described in the study, were “exactly according to the specification (Class A), 
throwing the elbows to the front (Class B), 
lifting the dumbbell only halfway (Class C), 
lowering the dumbbell only halfway (Class D) and 
throwing the hips to the front (Class E). 

By processing data gathered from accelerometers on the belt, forearm, arm, and dumbell of the participants in a machine learning algorithm, the question is can the appropriate activity quality (class A-E) be predicted?


```{r}
require("caret")
require("rpart")
require("nnet")
#require("rattle")#Causes my mac to crash
require("e1071")
require("gbm")
```

```{r}
train <- read.csv("/Users/Edward/Desktop/PracticalMachineLearning/pml-training.csv", header = TRUE, stringsAsFactors = FALSE)
```

We will view the dimension's of the training set

```{r}
dim(train)
```

We see there are 19622 observations of 160 variables.

```{r}
head(colnames(train))
```

#Data Preprocessing

We will be predicting the `classe`, but let us remove columns `X`, `user_name`, `raw_timestamp_part_1`, `raw_timestamp_part_2` and `cvtd_timestamp` as they should not be predictive in our model.

```{r}
train$X <- NULL
train$user_name <- NULL
train$raw_timestamp_part_1 <- NULL
train$raw_timestamp_part_2 <- NULL
train$cvtd_timestamp <- NULL
```


Let's partition data 80%-20%

```{r}
set.seed(123)
trainSet <- createDataPartition(y=train$classe, p=0.8, list=FALSE)
trainOne <- train[trainSet, ] 
testOne <- train[-trainSet, ]
```

We can remove variables with near-zero variance, which are unlikely to add predictive power to our model.

```{r}
lowVar <- nearZeroVar(trainOne, saveMetrics=TRUE)
head(lowVar)
```

Now we will remove these low-variability data from our sets:

```{r}
trainOne <- trainOne[,!(lowVar$nzv)]
testOne <- testOne[,!(lowVar$nzv)]
dim(trainOne)
```

Furthermore, throughout this analysis we will be ignoring NA values with `na.action = na.pass`.

#Modeling

Now we will generate an rpart model and evaluate it:

###CART (Classification and Regression Tree) Model

```{r}
attach(trainOne)
rpartModel <- train(classe ~ ., data = trainOne, method="rpart", na.action = na.pass)
rpartModel
```

Let's apply this model to our test set:

```{r}
confusionMatrix(predict(rpartModel, newdata=testOne, na.action = na.pass), testOne$classe)
```
###Out-of-sample Error

It is helpful to view the accuracy of the model:

```{r}
confusionMatrix(predict(rpartModel, newdata=testOne, na.action = na.pass), testOne$classe)$overall["Accuracy"]
```

Unfortunately, the accuracy is relatively low.

```{r}


```
